babette: BEAUti, BEAST and Tracer for R

### 2018-02-05 article.tex
EOD

\begin{lstlisting}[language=R, caption=Loading, label=lst:loading_babette, floatplacement=H]
library(babette)
\end{lstlisting}

@@

\begin{lstlisting}[language=R, caption=Example with different site model and clock model and tree prior, label=lst:all_different, floatplacement=H]
library(babette)
create_beast2_input_file(
  "alignment.fas",
  "beast2.xml",
  site_models = create_hky_site_model(),
  clock_models = create_rln_clock_model(),
  tree_priors = create_bd_tree_prior()
)
\end{lstlisting}

@@

\begin{lstlisting}[language=R, caption=Two alignments, label=lst:two_alignments, floatplacement=H]
library(babette)
babette::create_beast2_input_file(
  c("anthus_aco.fas", "anthus_nd2.fas"),
  "beast2.xml",
  site_models = list(
    create_tn93_site_model(), 
    create_gtr_site_model()
  )
)
\end{lstlisting}

@@

% TODO: Mix sentences better
Our initial motivation to create \verb;babette; 
is that we wanted to fix the crown age of a phylogeny, but because 
BEAUti assumes that a phylogeny has a crown age that needs to be jointly-estimated
with the phylogeny and other parameters. it does not allow for fixing
the crown age. Without \verb;babette;, one needs to manually edit the BEAST2 
XML configuration file, which is tedious and prone to errors. 
\verb;babette;, allows for easy setting of a fixed crown age, 
enabling theoretical experiments with one less source of variation. 
Setting a fixed crown age is not yet possible in BEAUti directly, 
but the BEAUti documentation mentions how to 
manually edit the XML file to allow for a fixed crown age. 
This is how to specify a fixed crown age with \verb;babette;:

@@

>>>>

\begin{lstlisting}[language=R, caption=Loading, label=lst:loading_babette, floatplacement=H]
library(babette)
library(beautier)
library(beastier)
library(tracerer)
\end{lstlisting}

@@

\begin{lstlisting}[language=R, caption=Example with different site model and clock model and tree prior, label=lst:all_different, floatplacement=H]
create_beast2_input_file(
  "alignment.fas",
  "beast2.xml",
  site_models = create_hky_site_model(),
  clock_models = create_rln_clock_model(),
  tree_priors = create_bd_tree_prior()
)
\end{lstlisting}

@@

\begin{lstlisting}[language=R, caption=Two alignments, label=lst:two_alignments, floatplacement=H]
create_beast2_input_file(
  c("anthus_aco.fas", "anthus_nd2.fas"),
  "beast2.xml",
  site_models = list(
    create_tn93_site_model(), 
    create_gtr_site_model()
  )
)
\end{lstlisting}

@@

% TODO: Mix sentences better
Our initial motivation to create \verb;babette; 
is that we wanted to fix the crown age of a phylogeny.
BEAUti assumes that a phylogeny has a crown age that needs to be jointly-estimated
with the phylogeny and other parameters. It does not allow for fixing
the crown age. Without \verb;babette;, one needs to manually edit the BEAST2 
XML configuration file, which is tedious and prone to errors. 
Fixing the crown ages allows for theoretical experiments 
with one less source of variation.

This is how to specify a fixed crown age with \verb;babette;:

@@

Instead of creating a BEAST2 input file and then running BEAST2,
\verb;babette; allows to do both steps at once:

\begin{lstlisting}[language=R, caption=Run BEAST2, label=lst:run_beast2, floatplacement=H]
out <- run_beast2(
  "alignment.fas"
)
\end{lstlisting}

\verb;run_beast2; also allows for setting a site model, clock model,
tree prior and a fixed crown age. The variable \verb;out; 
is a list that contains the posterior
parameter estimates, posterior phylogenies (one per alignment)
and operator acceptances.

\verb;babette; has functionality of both Tracer and
DensiTree. To calculate the effective sample sizes in \verb;babette;:

\begin{lstlisting}[language=R, caption=Calculate ESSes, label=lst:calc_esses, floatplacement=H]
traces <- remove_burn_ins(traces = out$estimates, burn_in_fraction = 0.2)
esses <- calc_esses(traces, sample_interval = 1000)
\end{lstlisting}

Tracer displays multiple summary statistics for each
estimated variable. To obtain, for example, the
summary statistics of the estimated birth rate in \verb;babette;:

sum_stats <- calc_sum_stats(
  out$estimates$birthRate, 
  sample_interval = 1000, burn_in_fraction = 0.1
)

DensiTree displays the phylogenies in a posterior.
To do the same in \verb;babette;:

\begin{lstlisting}[language=R, caption=Show posterior phylogenies, label=lst:densitree, floatplacement=H]
library(phangorn)
densiTree(out$anthus_aco_trees)
\end{lstlisting}
